---
title: "BayesC"
output:
  workflowr::wflow_html:
    toc: true
    latex_engine: "xelatex"
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

```{r 0-setup, include=FALSE, warning=FALSE}
#regular
if(1){
  library(dplyr)
  library(data.table)
  library(ggplot2)
  library(cowplot)
  library(qqman)
  library(viridis)
  library(scales)
  library(tidyverse)
  library(ggcorrplot)
  library(melt)
  library(reshape2)
  
  #options
  options(bitmapType = "cairo")
  options(error = function() traceback(3))
  
  #seed
  set.seed(123)
  
  #ggplot holder list
  gg <- vector(mode='list', length=12)
  #setwd('/data2/morgante_lab/nklimko/rep/dgrp-starve/')
}
```

```{r functions, warning=FALSE}
#reorders 50 X m table to a (50*m) X 2 table 
#converts method into factor to retain order
ggTidy <- function(data){
  
  for(i in 1:dim(data)[2]){
    
    name <- colnames(data)[i]
    temp <- cbind(rep(name, dim(data)[1]), data[,i, with=FALSE])
    
    if(i==1){
      hold <- temp
    } else{
      hold <- rbind(hold, temp, use.names=FALSE)
    }
  }
  colnames(hold) <- c("method", "cor")
  hold$method <- factor(hold$method, levels=unique(hold$method))
  
  
  return(hold)
}

#wrapper for ggplot call to custom fill sex, title, and y axis label
ggMake <- function(data, sex, custom.title, custom.Ylab){
  
  plothole <- ggplot(data, aes(x=method, y=cor, fill=method)) +
    geom_violin(color = NA, width = 0.65) +
    geom_boxplot(color='#440154FF', width = 0.15) +
    theme_minimal() +
    stat_summary(fun=mean, color='#440154FF', geom='point', 
                 shape=18, size=3, show.legend=FALSE) +
    labs(x=NULL,y=custom.Ylab, tag=sex, title=custom.title) +
    theme(legend.position='none',
          axis.text.x = element_text(angle = -45, size=10),
          text=element_text(size=10),
          plot.tag = element_text(size=15)) +
    scale_fill_viridis(begin = 0.4, end=0.9,discrete=TRUE)
  
  return(plothole)
  
}

```

```{r f-r2}
#bayesFinal <- readRDS("snake/data/zigure/bayesFinal.Rds")
b1 <- readRDS("snake/data/sr/34_opcor/f/bayesC_0.1.Rds")
b5 <- readRDS("snake/data/sr/34_opcor/f/bayesC_0.5.Rds")
b9 <- readRDS("snake/data/sr/34_opcor/f/bayesC_0.9.Rds")

bayesOp <- data.table(b1, b5, b9)
colnames(bayesOp) <- c('0.1', '0.5', '0.9')
data <- na.omit(ggTidy(bayesOp))

gg[[1]] <- ggMake(data, 'F', 'BayesC with Variable R2', 'Correlation')
```

```{r m-r2}

b1 <- readRDS("snake/data/sr/34_opcor/m/bayesC_0.1.Rds")
b5 <- readRDS("snake/data/sr/34_opcor/m/bayesC_0.5.Rds")
b9 <- readRDS("snake/data/sr/34_opcor/m/bayesC_0.9.Rds")

bayesOp <- data.table(b1, b5, b9)
colnames(bayesOp) <- c('0.1', '0.5', '0.9')
data <- na.omit(ggTidy(bayesOp))

gg[[2]] <- ggMake(data, 'M', 'BayesC with Variable R2', 'Correlation')

```

```{r f-trace}

base <- 'data/bglr/f/trace_'
base <- 'snake/data/bglr/f/trace_'
ETA <- paste0(base, 'ETA_1_parBayesC.dat')
varPath <- paste0(base, 'varE.dat')
fitPath <- paste0(base, 'fitBGLR')

fit <- readRDS(fitPath)

# Residual Variance
varE<-scan(varPath)
varFrame <- as.data.frame(varE)
colnames(varFrame) <- 'varVal'

gg[[3]] <- ggplot(varFrame, aes(x = 1:nrow(varFrame), y = varVal)) +
  labs(x=NULL,y='varE', tag='F', title='Residual Variance') +
  geom_line(linewidth=0.4) +
  geom_hline(yintercept = fit$varE, color='blue') +
  geom_vline(xintercept =  fit$burnIn/fit$thin, color='blue')


# Samples
#
TMP=read.table(ETA, header=T)
tmpFrame <- as.data.frame(TMP)
colnames(tmpFrame) <- 'tmp'

gg[[5]] <- ggplot(tmpFrame, aes(x = 1:nrow(tmpFrame), y = tmp)) +
  labs(x=NULL,y='TMP', tag='F', title='Probability of Inclusion') +
  geom_line(linewidth=0.4) +
  geom_hline(yintercept = fit$ETA[[1]]$probIn, color='blue') +
  geom_vline(xintercept =  fit$burnIn/fit$thin, color='blue')


```

```{r m-trace}

base <- 'data/bglr/m/trace_'
base <- 'snake/data/bglr/m/trace_'
ETA <- paste0(base, 'ETA_1_parBayesC.dat')
varPath <- paste0(base, 'varE.dat')
fitPath <- paste0(base, 'fitBGLR')

fit <- readRDS(fitPath)

# Residual Variance
varE<-scan(varPath)
varFrame <- as.data.frame(varE)
colnames(varFrame) <- 'varVal'

gg[[4]] <- ggplot(varFrame, aes(x = 1:nrow(varFrame), y = varVal)) +
  labs(x=NULL,y='varE', tag='M', title='Residual Variance') +
  geom_line(linewidth=0.4) +
  geom_hline(yintercept = fit$varE, color='blue') +
  geom_vline(xintercept =  fit$burnIn/fit$thin, color='blue')

# Samples
TMP=read.table(ETA, header=T)
tmpFrame <- as.data.frame(TMP)
colnames(tmpFrame) <- 'tmp'

gg[[6]] <- ggplot(tmpFrame, aes(x = 1:nrow(tmpFrame), y = tmp)) +
  labs(x=NULL,y='TMP', tag='M', title='Probability of Inclusion') +
  geom_line(linewidth=0.4) +
  geom_hline(yintercept = fit$ETA[[1]]$probIn, color='blue') +
  geom_vline(xintercept =  fit$burnIn/fit$thin, color='blue')

# Residual variance
#varE<-scan(varPath)
#plot(varE,type='o',col=2,cex=.5,ylab=expression(var[e]));
#abline(h=fit$varE,col=4,lwd=2)
#abline(v=fit$burnIn/fit$thin,col=4)

# Samples
#TMP=read.table(ETA, header=T)
#plot(TMP[,1],type='o',col=4, cex=0.5)
#abline(h=fit$ETA[[1]]$probIn,col=4,lwd=2)
#abline(v=fit$burnIn/fit$thin,col=4)

```

### Intro to Method

- BayesC - BayesC uses a spike-and-slab prior to perform effect shrinkage and variable selection. The model is fit using Markov Chain Monte Carlo(MCMC) methods.


### Parameters

For BayesC, we defined the following parameters:
- nIter : number of sampling iterations
- burnIn : iteration burn number, discarded from front end of sampling
- thin : factor to scale down nIter and burnIn for memory conservation
- R2 : proportion of variance explained by the model, 0 < R2 < 1


### Results


#### Proof of R2

R2 was chosen to be 0.9 after testing low, intermediate, and high values for R2. 

```{r plot-r9}

plot_grid(gg[[1]], gg[[2]], ncol=2)

```


#### Proof of niter/burnin/thin

nIter, burnIn, and thin were set to 130k, 30k, and 50 respectively. These parameters affect Markov Chain Monte Carlo sampling to generate a distribution of effect weights.

The following trace plots show the residual variance and samples for a single run:


##### Male

```{r plot-varE-f}

plot_grid(gg[[3]], gg[[5]], ncol=2)

```


##### Female

```{r plot-varE-m}

plot_grid(gg[[4]], gg[[6]], ncol=2)

```




