---
title: "Data Preparation"
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(data.table)
starve <- tibble(read.csv("data/starve.csv")) %>% select(-X)


```

# Phase 1
### [Full code]("../code/starveDataPrep.R")


Data was extracted from two tables of eQTL male/female traits. 

First, each file was filtered for non-null values

```{r qtl-readin, eval=FALSE}
  # csvPath dependent upon female/male file
  allRaw <- tibble(read.csv(csvPath))
  starveRaw <- allRaw %>% select(line,starvation) %>% filter(!is.na(starvation))
```

Two tables, starveF and starveM, were then joined one.
Difference and average of each line were calculated and added

```{r combineMF, eval=FALSE}
#Combine mf
starve <- starveF %>% mutate(m = starveM$m)

#compute difference and average
starve <- starve %>% select(line, f, m) %>% mutate(dif = f - m, avg = (f+m)/2)
```



# Phase 2
### [Full code]("../code/fourLinePrep.R")

Lines were manually selected and hard coded into the following extractions. 

```{r , eval=FALSE} 
#print(arrange(starve, starve$dif), n=205)
#print(arrange(starve, starve$avg), n=205)

# Lines with male SR high than female SR
difMinus <- starve %>% filter(dif < 0) %>% arrange(line) %>% select(line)
difMinus <- as.data.table(difMinus)
fwrite(difMinus, "./data/difMinus.txt")

# Lines with largest gap from male SR to female SR
difPlus <- starve %>% filter(dif >= 30) %>% arrange(line) %>% select(line)
difPlus <- as.data.table(difPlus)
fwrite(difPlus, "./data/difPlus.txt")

# Lowest average SR between males and females
avgMinus <- starve %>% filter(avg < 40) %>% arrange(line) %>% select(line)
avgMinus <- as.data.table(avgMinus)
fwrite(avgMinus, "./data/avgMinus.txt")

# Highest average SR between males and females
avgPlus <- starve %>% filter(dif >= 30) %>% arrange(line) %>% select(line)
avgPlus <- as.data.table(avgPlus)
fwrite(avgPlus, "./data/avgPlus.txt")

```


```{r dupeCheck, include=FALSE}
#dupeCheck <- rbind(difMinus, difPlus, avgMinus, avgPlus)
#print(arrange(dupeCheck, dupeCheck$line), n=55)
#8 of them repeat between dif and avg
#broken with data.table format

```

# Phase 3
### [Full code]("../code/combineSNP.R")

After creating the groups of interest, I 



