---
title: "Enrichment Overview"
output:
  workflowr::wflow_html:
    toc: true
    latex_engine: "xelatex"
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

```{r 0-setup, include=FALSE, warning=FALSE}
#regular
if(1){
  library(dplyr)
  library(data.table)
  library(ggplot2)
  library(cowplot)
  library(qqman)
  library(viridis)
  library(scales)
  library(tidyverse)
  library(ggcorrplot)
  library(melt)
  library(reshape2)
  library(knitr)
  library(kableExtra)
  library(DT)
  
  #options
  options(bitmapType = "cairo")
  options(error = function() traceback(3))
  
  #seed
  set.seed(123)
  
  #ggplot holder list
  gg <- vector(mode='list', length=12)
  
  
  options(knitr.kable.NA = '')
  
  
}
```



Dive back in: 

We left off looking to modify the selection cutoff for top terms. While various ideas were floated(standard devs, percentiles, flat cutoff), there was not a consistently most useful method for looking at both male and female data. As such, I opted to run a series of flat cutoffs(100, 50, 25) for both complete methods per sex.


I first looked at the top terms ordered to look for patterns in the data. While I'm unsure of the significance, the clustering of correlations for top terms shares a shape within sexes across methods. Notably, female data had a sharper increase(negative slope) for top results

```{r initialFinds}

load('snake/data/go/50_tables/enrichment.Rdata')

topGrid <- function(data, sex, psize, custom.title, custom.Xlab, custom.Ylab){
  plothole <- ggplot(data, aes(x=index, y=cor, label=term))+
    geom_point(color=viridis(1, begin=0.5), size=psize)+
    theme_minimal() +
    labs(x=custom.Xlab, y=custom.Ylab, tag=sex, title=custom.title) +
    theme(text=element_text(size=10), plot.tag = element_text(size=15))
  return(plothole)
}

gg[[1]] <- topGrid(blupF, 'F', 1, 'Top Female Results: TBLUP', 'Rank', 'Correlation')
gg[[2]] <- topGrid(blupM, 'M', 1, 'Top Male Results: TBLUP', 'Rank', 'Correlation')
gg[[3]] <- topGrid(bayesF, 'F', 1, 'Top Female Results: BayesC', 'Rank', 'Correlation')
gg[[4]] <- topGrid(bayesM, 'M', 1, 'Top Male Results: BayesC', 'Rank', 'Correlation')

plot_grid(gg[[1]], gg[[2]], gg[[3]], gg[[4]], ncol=2)

```

After this, I found the correlation between the two methods to see how similar generated results are. 


```{r foll}

print(cor(blupF$cor, bayesF$cor))

```

Female Overall

```{r ftop}

print(cor(blupF[1:200,cor], bayesF[1:200, cor]))

```

Female Top 200

```{r moll}

print(cor(blupM$cor, bayesM$cor))

```

Male Overall

```{r mtop}

print(cor(blupM[1:200,cor], bayesM[1:200, cor]))

```

Male Top 100


Moving past this, I wanted to assess the effect of term count on enrichment------000000------000000------000000------000000------000000




```{r dark}

load("snake/data/go/50_tables/enrich/kables.Rdata")

percentModder <- function(dataKable, custom.caption){
  
  dataKable[,5] <- dataKable[,5]*2
  dataKable[,8] <- dataKable[,8]*2
  
  colnames(dataKable) <- rep(c('Flybase Gene', 'Percent', 'Gene'), 3)

  kabled <- kable(dataKable, caption=custom.caption, "simple", header = c('Top 100 GO Terms' = 3, 'Top 50 GO Terms' = 3, 'Top 25 GO Terms' = 3))

  return(kabled)
  
}
bayesF_KableMod <- percentModder(bayesF_Kable, 'Female BayesC')

#print(bayesF_KableMod)

kable(bayesF_Kable, caption="Female BayesC", "simple")

if(0){
bayesF_KableMod <- percentModder(bayesF_Kable)
bayesM_KableMod <- percentModder(bayesM_Kable)
blupF_KableMod <- percentModder(blupF_Kable )
blupM_KableMod <- percentModder(blupM_Kable )

kable(bayesF_KableMod, caption="Female BayesC", "simple")
kable(bayesM_KableMod, caption="Male BayesC", "simple")
kable(blupF_KableMod, caption="Female TBLUP", "simple")
kable(blupM_KableMod, caption="Male TBLUP", "simple")
}

```

