---
title: "Linear Regression - Adjusted Phenotypes+"
output:
  workflowr::wflow_html:
    toc: true
    latex_engine: "xelatex"
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---
  
```{r 0-setup, include=FALSE, warning=FALSE}
library(dplyr)
library(data.table)
library(ggplot2)
library(cowplot)
library(qqman)
set.seed(1)

options(bitmapType = "cairo")

```

# Prep

```{r f-prep, evaluate=FALSE}

#load mgl data and custom function
load("/data/morgante_lab/data/dgrp/misc/adjustData.RData")

#read in expression data
f_data <- fread("data/fMeans.txt")

#grab starvation
y <- f_data[,starvation]

#store starvation with id
dat <- data.frame(id=f_data[,line], y=y)

#adjust data with given function
f_adjPheno <- adjustPheno(dat, "starvation")

#overwrite starvation column
f_adj <- f_data[,starvation:=f_adjPheno]

#save to txt file
fwrite(f_adj, "data/f_adj.txt")

```

```{r m-prep, evaluate=FALSE}

#load mgl data and custom function
load("/data/morgante_lab/data/dgrp/misc/adjustData.RData")

#read in expression data
m_data <- fread("data/mMeans.txt")

#grab starvation
y <- m_data[,starvation]

#store starvation with id
dat <- data.frame(id=m_data[,line], y=y)

#adjust data with given function
m_adjPheno <- adjustPheno(dat, "starvation")

#overwrite starvation column
m_adj <- m_data[,starvation:=m_adjPheno]

#save to txt file
fwrite(m_adj, "data/m_adj.txt")

```


# Correlations

In order to save time, I used secretariat's computational resources to perform the correlation calculations in parallel. The following two sections of code were submission scripts performed in secretariat.

```{r f-correlations, eval=FALSE}

library(dplyr)
library(data.table)
library(doParallel)

registerDoParallel(cores = 12)

# file path for final table
filePath <- "/data/morgante_lab/nklimko/rep/dgrp-starve/data/fRegress_adj.txt"

# gene expression data from earlier bound to line and starvation
f_adj <- fread("/data/morgante_lab/nklimko/rep/dgrp-starve/data/f_adj.txt")

start <- 3
end <- 11340
#f 11340
#m 13577

# foreach creates a list with index indicating position in loop
# function grabs the p-value from correlation of starvation to expression of trait
pvalList <- foreach(i=start:end) %dopar% {  
  temp <- summary(lm(f_adj[,c(2,i), with=FALSE], na.action=na.omit))[[4]][8]  
}

# converts list to vector
part <- unlist(pvalList, use.names = FALSE)

# binds p-vals to column index 
id <- start:end
part <- data.table(id,pvalList)

# write table to file
fwrite(part, filePath)

```

Both of these are available by clicking the drop down buttons on the right ->

```{r m-correlations, eval=FALSE}

library(dplyr)
library(data.table)
library(doParallel)

registerDoParallel(cores = 12)

# file path for final table
filePath <- "/data/morgante_lab/nklimko/rep/dgrp-starve/data/mRegress_adj.txt"

# gene expression data from earlier bound to line and starvation
m_adj <- fread("/data/morgante_lab/nklimko/rep/dgrp-starve/data/m_adj.txt")

start <- 3
end <- 13577
#f 11340
#m 13577

# foreach creates a list with index indicating position in loop
# function grabs the p-value from correlation of starvation to expression of trait
pvalList <- foreach(i=start:end) %dopar% {  
  temp <- summary(lm(m_adj[,c(2,i), with=FALSE], na.action=na.omit))[[4]][8]  
}

# converts list to vector
part <- unlist(pvalList, use.names = FALSE)

# binds p-vals to column index 
id <- start:end
part <- data.table(id,pvalList)

# write table to file
fwrite(part, filePath)

```

#### QQ and Manthattan Data Prep ->

```{r f-qq}

# Female qq prep

#f 11340
#m 13577

gg <- vector(mode='list', length=4)

#read in results from compute node
fReg <- fread("data/fRegress_adj.txt")

#number of data points
n <- dim(fReg)[1]

#Theoretical Quantiles
uniform <- 1:n/(n+1)

#sorted p values
empirical <- sort(fReg[,pvalList])

#-log 10
logPlot <- -log10(fReg[,pvalList])

#table for qqplot
qqdata <- data.table(uniform, empirical, logPlot)

qqdataF <- qqdata

# ggplot of data to fpr, qqplot
if(FALSE){
gg[[1]] <- ggplot(qqdata, aes(x=uniform, y=empirical)) +
  geom_point(color="red") +  
  geom_segment(aes(x=0,y=0,xend=1,yend=1)) + 
  xlab("Theoretical Quantiles") +
  ylab("Sorted p-values") +
  ggtitle("Female Expression QQ Plot")
}

gg[[2]] <- ggplot(qqdata, aes(x=uniform, y=logPlot)) +
  geom_point(color="red") + 
  geom_hline(yintercept= -log10(0.05/n), color="magenta") +
  xlab("Theoretical Quantiles") +
  ylab("-log10 p-values") +
  ggtitle("Female Expression Manhattan Plot")



## Code for looking at specific genes
#fMeans <- fread("data/fMeans.txt")
#genes <- colnames(fMeans)
#genes <- genes[3:length(genes)]
#fReg <- fReg[, gene:=genes]

```

```{r m-qq}

# Male qq prep

#read in results from compute node
mReg <- fread("data/mRegress_adj.txt")

#number of data points
n <- dim(mReg)[1]

#Theoretical Quantiles
uniform <- 1:n/(n+1)

#sorted p values
empirical <- sort(mReg[,pvalList])

#-log 10
logPlot <- -log10(mReg[,pvalList])

#table for qqplot
qqdata <- data.table(uniform, empirical, logPlot)

qqdataM <- qqdata


# ggplot of data to fpr, qqplot
if(FALSE){
gg[[3]] <- ggplot(qqdata, aes(x=uniform, y=empirical)) +
  geom_point(color="blue") +  
  geom_segment(aes(x=0,y=0,xend=1,yend=1), linewidth=1) + 
  xlab("Theoretical Quantiles") +
  ylab("Sorted p-values") +
  ggtitle("Male Expression QQ Plot")
}


gg[[4]] <- ggplot(qqdata, aes(x=uniform, y=logPlot)) +
  geom_point(color="blue") + 
  geom_hline(yintercept= -log10(0.05/n), color="cyan") +
  xlab("Theoretical Quantiles") +
  ylab("-log10 p-values") +
  ggtitle("Male Expression Manhattan Plot")

## Code for looking at specific genes
#mMeans <- fread("data/mMeans.txt")
#genes <- colnames(mMeans)
#genes <- genes[3:length(genes)]
#mReg <- mReg[, gene:=genes]

```

# QQ Plots

```{r qq-plot, warning=FALSE}

par(mfrow=c(1,2))

qq(qqdataF[,empirical], main="Female Gene p-values")
qq(qqdataM[,empirical], main="Male Gene p-values")

```

# Manhattan Plots

```{r manhattan-plot, warning=FALSE}

par(mfrow=c(1,2))

plot_grid(gg[[2]],gg[[4]], ncol=2)

```

Part of the reason both plots deviate from linearity because we did not account for genes that were correlated with each other.


