---
title: "Simple Regression"
output:
  workflowr::wflow_html:
    toc: true
    latex_engine: "xelatex"
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---
  
```{r 0-setup, include=FALSE, warning=FALSE}
library(dplyr)
library(data.table)
library(ggplot2)
library(cowplot)
library(qqman)
set.seed(1)

options(bitmapType = "cairo")

```

# Prep

```{r f-prep, evaluate=FALSE}

# The following was run to prepare a data table for job submission

#expression data
xpf <- fread("data/xp-f.txt")

#line/trait data
dtf <- fread("data/eQTL_traits_females.csv")

#select line and starvation only, remove nulls
dtf <- na.omit(dtf[,c(1,10)])

fwrite(dtf, "data/starve-f.txt")

fMeans <- dtf[xpf, on = .(line), nomatch=NULL, all=TRUE]

fwrite(fMeans, "data/fMeans.txt")
```

```{r m-prep, evaluate=FALSE}

# The following was run to prepare a data table for job submission

#expression data
xpm <- fread("data/xp-m.txt")

#line/trait data
dtm <- fread("data/eQTL_traits_males.csv")

#select line and starvation only, remove nulls
dtm <- na.omit(dtm[,c(1,11)])

fwrite(dtm, "data/starve-m.txt")

mMeans <- dtm[xpm, on = .(line), nomatch=NULL, all=TRUE]

fwrite(mMeans, "data/mMeans.txt")
```

# Demo

As a demo, I've walked through the process for finding a correlation for the first gene. Below is a plot of starvation vs flybase gene 3:

```{r process}

plot(mMeans[,c(3,2), with=FALSE])

```

lm() returns the simple regression of y to x, giving parameters for a slope-intercept form equation

```{r process-lm}

y <- mMeans[,starvation]
x <- mMeans[,3,with=FALSE]
lm(formula=unlist(y)~unlist(x), na.action=na.omit)
```

Plotting this line shows the effect of the particular gene(x) on starvation resistance(y). 

```{r process-abline}

plot(mMeans[,c(3,2), with=FALSE])
abline(22.91, 5.01)

```

summary() returns a variety of useful statistics regarding simple regression. 
```{r process-summary}
summary(lm(mMeans[,c(2,3), with=FALSE], na.action=na.omit))
```

The p-value is what we're interested in, and can be accessed from the summary object at [[4]][8]

The following line of code combine all of these parts into one to extract the p-value from the correlation between starvation resistance and any gene denoted by column i. 

```{r product, eval=FALSE}
summary(lm(mMeans[,c(2,3), with=FALSE], na.action=na.omit))[[4]][8]
```

This will need to be run over 24 thousand times to capture every gene expressed in both females and males. 

# Correlations

In order to save time, I used secretariat's computational resources to perform the correlation calculations in parallel. The following two sections of code were submission scripts performed in secretariat.

```{r f-correlations, eval=FALSE}

library(dplyr)
library(data.table)
library(doParallel)

registerDoParallel(cores = 12)

# file path for final table
filePath <- "/data/morgante_lab/nklimko/rep/dgrp-starve/data/fRegress.txt"

# gene expression data from earlier bound to line and starvation
fMeans <- fread("/data/morgante_lab/nklimko/rep/dgrp-starve/data/fMeans.txt")

start <- 3
end <- 11340
#f 11340
#m 13577

# foreach creates a list with index indicating position in loop
# function grabs the p-value from correlation of starvation to expression of trait
pvalList <- foreach(i=start:end) %dopar% {  
  temp <- summary(lm(fMeans[,c(2,i), with=FALSE], na.action=na.omit))[[4]][8]  
}

# converts list to vector
part <- unlist(pvalList, use.names = FALSE)

# binds p-vals to column index 
id <- start:end
part <- data.table(id,pvalList)

# write table to file
fwrite(part, filePath)

```

Both of these are available by clicking the drop down buttons on the right ->

```{r m-correlations, eval=FALSE}
library(dplyr)
library(data.table)
library(doParallel)

registerDoParallel(cores = 12)

# file path for final table
filePath <- "/data/morgante_lab/nklimko/rep/dgrp-starve/data/mRegress.txt"

# gene expression data from earlier bound to line and starvation
mMeans <- fread("/data/morgante_lab/nklimko/rep/dgrp-starve/data/mMeans.txt")

start <- 3
end <- 13577

# foreach creates a list with index indicating position in loop
# function grabs the p-value from correlation of starvation to expression of trait
pvalList <- foreach(i=start:end) %dopar% {
  
  temp <- summary(lm(mMeans[,c(2,i), with=FALSE], na.action=na.omit))[[4]][8]  
}

# converts list to vector
part <- unlist(pvalList, use.names = FALSE)

# binds p-vals to column index 
id <- start:end
part <- data.table(id,pvalList)

#writes table to file
fwrite(part, filePath)

```

#### QQ and Manthattan Data Prep ->

```{r f-qq}

# Female qq prep

#f 11340
#m 13577

gg <- vector(mode='list', length=4)

#read in results from compute node
fReg <- fread("data/fRegress.txt")

#number of data points
n <- dim(fReg)[1]

#uniform distribution
uniform <- 1:n/(n+1)

#sorted p values
empirical <- sort(fReg[,pvalList])

#-log 10
logPlot <- -log10(fReg[,pvalList])

#table for qqplot
qqdata <- data.table(uniform, empirical, logPlot)

qqdataF <- qqdata

# ggplot of data to fpr, qqplot
if(FALSE){
gg[[1]] <- ggplot(qqdata, aes(x=uniform, y=empirical)) +
  geom_point(color="red") +  
  geom_segment(aes(x=0,y=0,xend=1,yend=1)) + 
  xlab("Uniform Distribution") +
  ylab("Sorted p-values") +
  ggtitle("Female Expression QQ Plot")
}

gg[[2]] <- ggplot(qqdata, aes(x=uniform, y=logPlot)) +
  geom_point(color="red") + 
  geom_hline(yintercept= -log10(0.05/n), color="magenta") +
  xlab("Uniform Distribution") +
  ylab("-log10 p-values") +
  ggtitle("Female Expression Manhattan Plot")



## Code for looking at specific genes
#fMeans <- fread("data/fMeans.txt")
#genes <- colnames(fMeans)
#genes <- genes[3:length(genes)]
#fReg <- fReg[, gene:=genes]

```

```{r m-qq}

# Male qq prep

#read in results from compute node
mReg <- fread("data/mRegress.txt")

#number of data points
n <- dim(mReg)[1]

#uniform distribution
uniform <- 1:n/(n+1)

#sorted p values
empirical <- sort(mReg[,pvalList])

#-log 10
logPlot <- -log10(mReg[,pvalList])

#table for qqplot
qqdata <- data.table(uniform, empirical, logPlot)

qqdataM <- qqdata


# ggplot of data to fpr, qqplot
if(FALSE){
gg[[3]] <- ggplot(qqdata, aes(x=uniform, y=empirical)) +
  geom_point(color="blue") +  
  geom_segment(aes(x=0,y=0,xend=1,yend=1), linewidth=1) + 
  xlab("Uniform Distribution") +
  ylab("Sorted p-values") +
  ggtitle("Male Expression QQ Plot")
}


gg[[4]] <- ggplot(qqdata, aes(x=uniform, y=logPlot)) +
  geom_point(color="blue") + 
  geom_hline(yintercept= -log10(0.05/n), color="cyan") +
  xlab("Uniform Distribution") +
  ylab("-log10 p-values") +
  ggtitle("Male Expression Manhattan Plot")

## Code for looking at specific genes
#mMeans <- fread("data/mMeans.txt")
#genes <- colnames(mMeans)
#genes <- genes[3:length(genes)]
#mReg <- mReg[, gene:=genes]

```

# QQ Plots

```{r qq-plot, warning=FALSE}

par(mfrow=c(1,2))

qq(qqdataF[,empirical], main="Female Gene p-values")
qq(qqdataM[,empirical], main="Male Gene p-values")

```

# Manhattan Plots

```{r manhattan-plot, warning=FALSE}

par(mfrow=c(1,2))

plot_grid(gg[[2]],gg[[4]], ncol=2)

```

Part of the reason both plots deviate from linearity because we did not account for genes that were correlated with each other.


