---
title: "PCA with some scatter plots"
output:
  workflowr::wflow_html:
    toc: true
    latex_engine: "xelatex"
    code_folding: "hide"
editor_options:
  chunk_output_type: console
---

```{r 0-setup, include=FALSE, echo=FALSE, warning=FALSE}
library(dplyr)
library(stringr)
library(data.table)
library(ggplot2)
library(cowplot)

#gg[[1]],gg[[2]],gg[[3]],gg[[4]],gg[[5]],gg[[6]],gg[[7]],gg[[8]],gg[[9]],gg[[10]],gg[[11]],gg[[12]],gg[[13]],gg[[14]],gg[[15]],gg[[16]],gg[[17]]

```


# Multiple Comparison

Starvation was plotted against every other trait and correlations were calculated. Ideally, strong correlation between traits can be used to increase the impact of PCA as traits with strong correlations can be used to reduce dimensions. Strong correlations allow for calculated columns.

```{r multicomp}

# Read in tables
dtf <- fread("data/eQTL_traits_females.csv")
dtm <- fread("data/eQTL_traits_males.csv")

#Change column order to line, starvation, everything else
setcolorder(dtf, c(1,10,2:9,11:19))
#male has aggression moved to end to line up all other column indices
setcolorder(dtm, c(1,11,3:10,12:20,2))

# Bind tables on line
bound <- dtm[dtf, on=.(line)]

# create average starvation
x <- (bound[,2,with=FALSE] + bound[,21,with=FALSE]) / 2

storage <- rep(0,18)
gg <- vector(mode='list', length=18)

  # AGGRESSION, male only


yL <- colnames(bound[, 20, with=FALSE])

x <- bound[,2] 
y <- bound[,20]
raw <- na.omit(cbind(x,y))
clean <- setDT(raw)
colnames(clean) <- c("starvation", "trait")

corCoeff <- str_sub(as.character(cor.test(clean$starvation, clean$trait)[4]), 8,16)
storage[1] <- as.numeric(corCoeff)

graphTitle <- paste0(colnames(bound[, 20, with=FALSE]),": R=",corCoeff)

gg[[1]] <- ggplot(clean, aes(x=starvation, y=trait)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm) +
  labs(title = graphTitle, x="starvation",y=yL)

# Determine average of each trait
for(i in 3:19)
{
  #title <- paste("starvation resistance vs",colnames(bound[, i, with=FALSE]))
  j <- i + 19
  y <- (bound[,i,with=FALSE] + bound[,j,with=FALSE]) / 2


yL <- colnames(bound[, i, with=FALSE])

raw <- na.omit(cbind(x,y))
clean <- setDT(raw)
colnames(clean) <- c("starvation", "trait")


corCoeff <- str_sub(as.character(cor.test(clean$starvation, clean$trait)[4]), 8,16)
storage[i-1] <- as.numeric(corCoeff)

graphTitle <- paste0(colnames(bound[, i, with=FALSE]),": R=",corCoeff)

gg[[i-1]] <- ggplot(clean, aes(x=starvation, y=trait)) +
  geom_point() +
  geom_smooth(formula = y ~ x, method=lm) +
  labs(title = graphTitle, x="starvation",y=yL)
}


gg <- cbind(gg,storage)

gg <- gg[order(abs(storage),decreasing=TRUE),]

```


```{r plots, echo=FALSE}

par(mfrow=c(2,2))
plot_grid(gg[[1]],gg[[2]],gg[[3]],gg[[4]],ncol=2)
plot_grid(gg[[5]],gg[[6]],gg[[7]],gg[[8]],ncol=2)
plot_grid(gg[[9]],gg[[10]],gg[[11]],gg[[12]],ncol=2)
plot_grid(gg[[13]],gg[[14]],gg[[15]],gg[[16]],ncol=2)
par(mfrow=c(2,1))
plot_grid(gg[[17]],gg[[18]])



```


# Principle Component Analyses

The goal of PCA with any dataset is to reduce the dimensionality into simpler and more intuitive bases. 



#### Useful Links

[Thorough explanation of PCA](https://youtu.be/FgakZw6K1QQ)

[PCA in R](https://www.youtube.com/watch?v=0Jp4gsfOLMs)

[prcomp function](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/prcomp)

## Combined

This approach will mostly capture the difference between male and female expression levels with the intention of revealing starvatino resistance patterns within each subgroup.

```{r combinedPCA}

dtf <- fread("data/eQTL_traits_females.csv")
dtm <- fread("data/eQTL_traits_males.csv")
xpf <- fread("data/xp-f.txt")
xpm <- fread("data/xp-m.txt")

#extract starvation and line
lsf <- dtf[,c(1,10)]
lsm <- dtm[,c(1,11)]

testF <- merge(xpf,lsf,all.x=TRUE)
testM <- merge(xpm,lsm,all.x=TRUE)

testF[,group:="female"]
testM[,group:="male"]

#move starvation label to front, after line
setcolorder(testF, c(1,11340:11341,2:11339))
setcolorder(testM, c(1,13577:13578,2:13576))

xpa <- rbind(testF, testM, fill=TRUE)

#set all NULL to zero
xpa[is.na(xpa)] <- 0


#PCA analysis
pcs <- prcomp(xpa[,c(-1,-2,-3)], scale=TRUE)

#compute variance and normalize to 100%
pcVar <- pcs$sdev^2
pcVarNorm <- round(pcVar/sum(pcVar)*100,2)

#Scree plot data of first 10 PCs
scree <- data.frame(1:10, pcVarNorm[1:10])
colnames(scree) <- c("PC", "variance")
 
#ggplot arguments - Scree plot
ggplot(data=scree, aes(x=PC, y=variance)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=variance), vjust=-0.3, size=3.5) +
  ggtitle("Combined Scree Plot")

#plot data from PCA to ggplot
pcaPlot <- data.frame(xpa[,1:3], X=pcs$x[,1], Y=pcs$x[,2])

#Group identification
ggplot(data=pcaPlot, aes(x=X, y=Y, color=group)) +
  geom_point() +
  xlab(paste("PC1: ",pcVarNorm[1],"%", sep="")) +
  ylab(paste("PC2: ",pcVarNorm[2],"%", sep="")) +
  ggtitle("Combined PCA")

#starvation patterns
ggplot(data=pcaPlot, aes(x=X, y=Y, color=starvation)) +
  geom_point() +
  scale_color_gradient(low="yellow", high="red") +
  xlab(paste("PC1: ",pcVarNorm[1],"%", sep="")) +
  ylab(paste("PC2: ",pcVarNorm[2],"%", sep="")) +
  ggtitle("Combined PCA")

  


genCor <- cor(xpa[,c(-1,-2,-3)], pcaPlot[,4:5])

genCor <- as.data.frame(genCor)


corPC1 <- arrange(genCor,desc(abs(X)))
corPC2 <- arrange(genCor,desc(abs(Y)))

corPC1[1:25,1:2]
corPC2[1:25,1:2]



```


## Female

Looking at female expression alone against female starvation resistance to identify patterns in population structure.

```{r femalePCA}

# Read in data
xpf <- fread("data/xp-f.txt")

xpa <- xpf

#store lines, remove, convert to matrix, and label rows
xpaLines <- xpa[,line]
xpa <- xpa[,line:=NULL]
xpa <- as.matrix(xpa)
row.names(xpa) <- xpaLines

# performs PCA, scale=TRUE autocenters data around means
pcs <- prcomp(xpa, scale=TRUE)

#compute variance and normalize to 100%
pcVar <- pcs$sdev^2
pcVarNorm <- round(pcVar/sum(pcVar)*100,2)

#Scree plot data of first 10 PCs
scree <- data.frame(1:10, pcVarNorm[1:10])
colnames(scree) <- c("PC", "variance")

#ggplot arguments - Scree plot
ggplot(data=scree, aes(x=PC, y=variance)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=variance), vjust=-0.3, size=3.5) +
  ggtitle("Female Scree Plot")


#PCA data
pcaPlot <- data.frame(line=rownames(pcs$x), pcs$x[,1], pcs$x[,2])
starveLines <- as.data.frame(dtf[,1:2])
pcaPlot <- na.omit(merge(x=pcaPlot,y=starveLines, by="line"))
colnames(pcaPlot) <- c("line","X","Y","starve")


#ggplot arguments - PC plot
ggplot(data=pcaPlot, aes(x=X, y=Y, label=line, color=starve)) +
  geom_point() +
  scale_color_gradient(low="yellow", high="red") +
  xlab(paste("PC1: ",pcVarNorm[1],"%", sep="")) +
  ylab(paste("PC2: ",pcVarNorm[2],"%", sep="")) +
  ggtitle("Female PCA")
  


```

Quantitative approach: correlations for the top 25 genes for each principal component.

```{r}

genCor <- cor(xpa, pcs$x[,1:2])

genCor <- as.data.frame(genCor)

corPC1 <- arrange(genCor,desc(abs(PC1)))
corPC2 <- arrange(genCor,desc(abs(PC2)))

print("Top 25 genes for PC1:")
corPC1[1:25,1:2]

print("Top 25 genes for PC2")
corPC2[1:25,1:2]



```

## Male

Looking at male expression alone against male starvation resistance to identify patterns in population structure.

```{r malePCA}
# Read in data
xpm <- fread("data/xp-m.txt")

xpa <- xpm


#store lines, remove, convert to matrix, and label rows
xpaLines <- xpa[,line]
xpa <- xpa[,line:=NULL]
xpa <- as.matrix(xpa)
row.names(xpa) <- xpaLines

# performs PCA, scale=TRUE autocenters data around means
pcs <- prcomp(xpa, scale=TRUE)

#compute variance and normalize to 100%
pcVar <- pcs$sdev^2
pcVarNorm <- round(pcVar/sum(pcVar)*100,2)

#Scree plot data of first 10 PCs
scree <- data.frame(1:10, pcVarNorm[1:10])
colnames(scree) <- c("PC", "variance")

#ggplot arguments - Scree plot
ggplot(data=scree, aes(x=PC, y=variance)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=variance), vjust=-0.3, size=3.5) +
  ggtitle("Male Scree Plot")


#PCA data
pcaPlot <- data.frame(line=rownames(pcs$x), pcs$x[,1], pcs$x[,2])
starveLines <- as.data.frame(dtf[,1:2])
pcaPlot <- na.omit(merge(x=pcaPlot,y=starveLines, by="line"))
colnames(pcaPlot) <- c("line","X","Y","starve")


#ggplot arguments - PC plot
ggplot(data=pcaPlot, aes(x=X, y=Y, label=line, color=starve)) +
  geom_point() +
  scale_color_gradient(low="yellow", high="red") +
  xlab(paste("PC1: ",pcVarNorm[1],"%", sep="")) +
  ylab(paste("PC2: ",pcVarNorm[2],"%", sep="")) +
  ggtitle("Male PCA")
  


```

Quantitative approach: correlations for the top 25 genes for each principal component.

```{r}

genCor <- cor(xpa, pcs$x[,1:2])

genCor <- as.data.frame(genCor)

corPC1 <- arrange(genCor,desc(abs(PC1)))
corPC2 <- arrange(genCor,desc(abs(PC2)))

print("Top 25 genes for PC1:")
corPC1[1:25,1:2]

print("Top 25 genes for PC2")
corPC2[1:25,1:2]



```




## Troubleshooting

I plotted the top ten principal components against each other to visally identify any clustering among any of the components. all principal components fail to separate cleanly. 



```{r groupingcheck}

# Read in data
xpm <- fread("data/xp-m.txt")

xpa <- xpm


#store lines, remove, convert to matrix, and label rows
xpaLines <- xpa[,line]
xpa <- xpa[,line:=NULL]
xpa <- as.matrix(xpa)
row.names(xpa) <- xpaLines

# performs PCA, scale=TRUE autocenters data around means
pcs <- prcomp(xpa, scale=TRUE)

#compute variance and normalize to 100%
pcVar <- pcs$sdev^2
pcVarNorm <- round(pcVar/sum(pcVar)*100,2)

#PCA data
pcaPlot <- data.frame(pcs$x[,1:10])

par(mfrow=c(2,3))

for(i in 1:10){
  for(j in 1:10){
    if(i!=j){
      plot(pcaPlot[,i], pcaPlot[,j])
      
    }else{
      #nothing
    }
  }
}

#ggplot arguments - PC plot







```







